Is high social class always beneficial for survival? Northern Sweden 1801--2013

### 2018-03-15 CreData/fromFobLisa.Rmd
fix Lexis diagram

This is an updated version of *showData4.Rmd*, where we used the newly-found
knowledge that the yearly *lisa* data relates to *December 31* the given
year, while the *FoB* data relate to the date *November 1* the given
year. These facts have the following implications for our plan: For *FoB*
data, we follow individuals from their appearance in a FoB until the next
FoB five years later, even if they do not appear there. The exception is if
they die within five years after the first appearance. So, *fob60* induces
a time interval *from November 1, 1960 to October 31, 1965*, and so on. 
Since we do not have birth dates but only birth year and quarter,
we decide to use as birthdate the *midpoint of the birth quarter*. 
WE work with all ages 0-110, and only as the final step, `age.window` is called.

@@

```{r styronper}
load("../../Data/socstyr.rda")
indx <- match(per$LINNEID, socstyr$LINNEID)
per$DODSDAT <- socstyr$DODSDAT[indx]
fixd <- !is.na(per$DODSDAT) & (per$DODSDAT < 1960010)
per$DODSDAT[fixd] <- per$DODSDAT[fixd] * 100 + 15
per$deathdate <- as.numeric(skum::tillTid(per$DODSDAT))
per$dodsKOMMUN <- socstyr$KOMMUN[indx]
per$dodsALDER <- socstyr$ALDER[indx]
per$ULORSAK <- socstyr$ULORSAK[indx]
```

>>>>

This is an updated version of *showData4.Rmd*, where we used the newly-found
knowledge that the yearly *lisa* data relates to *December 31* the given
year, while the *FoB* data relate to the date *November 1* the given
year. These facts have the following implications for our plan: For *FoB*
data, we follow individuals from their appearance in a FoB until the next
FoB five years later, even if they do not appear there. The exception is if
they die within five years after the first appearance. So, *fob60* induces
a time interval *from November 1, 1960 to October 31, 1965*, and so on. 
Since we do not have birth dates but only birth year and quarter,
we decide to use as birthdate the *midpoint of the birth quarter*. 
We work with all ages 0-110, and only as the final step, `age.window` is called.

@@

```{r styronper}
load("../../Data/socstyr.rda")
indx <- match(per$LINNEID, socstyr$LINNEID)
per$DODSDAT <- socstyr$DODSDAT[indx]
fixd <- !is.na(per$DODSDAT) & (per$DODSDAT < 1960010) # Only month and year
per$DODSDAT[fixd] <- per$DODSDAT[fixd] * 100 + 15
per$deathdate <- as.numeric(skum::tillTid(per$DODSDAT))
per$dodsKOMMUN <- socstyr$KOMMUN[indx]
per$dodsALDER <- socstyr$ALDER[indx]
per$ULORSAK <- socstyr$ULORSAK[indx]
```

### 2018-03-15 marginal.Rmd
fix Lexis diagram

```{r lexis2,fig.cap="The sampling frame.", fig.height = 5}
plot(c(1801, 1801, 1951, 1951, 1801), c(30, 100, 100, 30, 30),
     type = "l", xlim = c(1800, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 100))
axis(1, at = c(1801, 1851, 1901, 1951, 1976, 1986, 1996, 2005, 2014), las = 2)
axis(2, las = 2, at = c(0, 30, 65, 100))
axis(4, las = 2, at = c(0, 30, 65, 100))
box()
abline(h = 0)
lines(c(1976, 1976, 2014, 2014, 1976), c(30, 100, 100, 30, 30),
      col = "blue", lwd = 2)
#
pn.exp <- round(sum(skume1$exit - skume1$enter) + sum(skume2$exit - skume2$enter))
pn.d <-  as.integer(sum(skume1$event) + sum(skume2$event))
#
text(1875, 70, "PopLink", cex = 2, col = "red")
##text(1899, 60, paste(pn.rec, "records"))
text(1875, 55, paste(round(pn.exp / 1000000, 1), "million p-years"))
text(1874.5, 50, paste(round(pn.d / 1000, 1), "thousand deaths"))
#
n.exp <- round(sum(foli1$exit - foli1$enter) + sum(foli2$exit -foli2$enter))
n.d <-  as.integer(sum(foli1$event) + sum(foli2$event))

>>>>

```{r lexis2,fig.cap="The sampling frame.", fig.height = 5}
plot(c(1801, 1801, 1951, 1951, 1801), c(40, 100, 100, 40, 40),
     type = "l", xlim = c(1800, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 100))
axis(1, at = c(1801, 1851, 1901, 1951, 1976, 1986, 1996, 2005, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 100))
axis(4, las = 2, at = c(0, 40, 65, 100))
box()
abline(h = 0)
lines(c(1976, 1976, 2014, 2014, 1976), c(40, 100, 100, 40, 40),
      col = "blue", lwd = 2)
#
pn.exp <- round(sum(skume1$exit - skume1$enter) + sum(skume2$exit - skume2$enter))
pn.d <-  as.integer(sum(skume1$event) + sum(skume2$event))
#
text(1875, 70, "PopLink", cex = 2, col = "red")
##text(1899, 60, paste(pn.rec, "records"))
text(1875, 55, paste(round(pn.exp / 1000000, 1), "million p-years"))
text(1874.5, 50, paste(round(pn.d / 1000, 1), "thousand deaths"))
#
n.exp <- round(sum(foli1$exit - foli1$enter) + sum(foli2$exit -foli2$enter))
n.d <-  as.integer(sum(foli1$event) + sum(foli2$event))
